Advantages of mixed quantum gates for information processing

### 2019-03-20 decorrelation/mixed_unitaries.tex
added in some references, now I need to sleep

The ultimate impact of a gate error on a quantum circuit depends strongly on both the magnitude and the nature of the error. Systematic, or \emph{coherent}, errors can arise from poorly calibrated controls or imperfect gate compilations that induce repeatable, undesired unitary errors on the state of a QIP. Errors of this type are correlated in time may add up constructively or destructively, depending on the circuit. They are are  computationally expensive to model and it can be difficult to place tight analytic bounds on circuit performance [cite]. Contrast this against random, or \emph{stochastic}, errors, which often result from high-frequency noise in the controls or the environment. Systems with stochastic errors can usually be modeled by defining a rate of various discrete errors in the system, such as a bit flips or phase flips. These errors are significantly easier to simulate on a classical computer, and their impact on quantum circuits is much easier to estimate [Ken Brown, Steve Flammia].

@@

\subsection{Representing CPTP maps}
\noindent CPTP maps possesses a number of useful representations, including Kraus operators[], Choi matrices[], and Jamiolkowsi states[]. But for the purposes of this article, the \emph{process matrix} representation will be particularly convenient [], and we shall denote the process matrix associated with a given CPTP map $\mathsf{G}$ with the corresponding calligraphic character, $\mathcal{G}$. A key feature of process matrices is that they act through the usual matrix multiplication on the vectorized quantum state:
\begin{align}
	\vectorize{\mathsf{\tilde G}(\rho)}
		&= \mathcal{\tilde G}\cdot\vectorize{\rho} \\
		&= \mathcal{E}\cdot\mathcal{G}\cdot\vectorize{\rho}
\end{align}
We shall work in the basis Pauli matrices, defining $\Sigma = \left\{I, \sigma_x, \sigma_y, \sigma_z\right\}^{\otimes n}$ as the collection of all $4^n$ $n$-qubit Pauli operators.  In this basis,
\begin{equation}
  \vectorize{\rho}_i = \tr\!\left(\rho \cdot\Sigma_i\right)/2^n,
\end{equation}
and
\begin{equation}
	(\mathcal{\tilde G})_{ij} = \tr\!\left(\mathsf{\tilde G}\left(\Sigma_j\right) \cdot\Sigma_i\right)/2^n.
\end{equation}
Process matrices written in Pauli basis are is often referred to as a \emph{Pauli transfer matrix} []. Error maps take an especially nice form in this representation:
\begin{equation}\label{eq:process_matrix}
\mathcal{E} =
	\left(\begin{array}{c|cccc}
		1 &  & \vec{0}^T & \\ 
		\hline & &  &  \\
		\vec{m} &  & R &  \\
		 &  &  & 
	\end{array} 	
	\right)
\end{equation}
The top row of all trace-preserving (TP) maps is fixed to $\{1,0,0,0,\cdots\}$ and the remainder of the first column, $\vec{m}$, describes any deviations from unitality, as could arise from amplitude damping. If the error map is unitary, then the error is coherent, and the unital submatrix $R$ is perfectly antisymmetric, corresponding to a rotation of the generalized Bloch vector. Importantly, if $R$ is diagonal, then the error  is Pauli stochastic, with each diagonal entry corresponding to the probability that its associated Pauli error occurs in each application of the gate. 
% If R is symmetric but not diagonal, then the channel is still stochastic, but the random errors consist of correlated Pauli operators (such as $X+Y$). For a single qubit, this describes everything, but the situation can be slightly more complicated for more qubits.

>>>>

The ultimate impact of a gate error on a quantum circuit depends strongly on both the magnitude and the nature of the error. Systematic, or \emph{coherent}, errors can arise from poorly calibrated controls or imperfect gate compilations that induce repeatable, undesired unitary errors on the state of a QIP. Errors of this type are correlated in time may add up constructively or destructively, depending on the circuit. They are are  computationally expensive to model and it can be difficult to place tight analytic bounds on circuit performance [cite]. Contrast this against random, or \emph{stochastic}, errors, which often result from high-frequency noise in the controls or the environment. Systems with stochastic errors can usually be modeled by defining a rate of various discrete errors in the system, such as a bit flips or phase flips. These errors are significantly easier to simulate on a classical computer, and their impact on quantum circuits is much easier to estimate \cite{Beale2018}.

@@

\subsection{Representing CPTP maps}
\noindent CPTP maps possesses a number of useful representations, including Kraus operators\cite{1983}, Choi matrices\cite{Choi1975}, and Jamiolkowsi states\cite{yczkowski2004}. But for the purposes of this article, the \emph{process matrix} representation will be particularly convenient\cite{OBrien2004}, and we shall denote the process matrix associated with a given CPTP map $\mathsf{G}$ with the corresponding calligraphic character, $\mathcal{G}$. A key feature of process matrices is that they act through the usual matrix multiplication on the vectorized quantum state:
\begin{align}
	\vectorize{\mathsf{\tilde G}(\rho)}
		&= \mathcal{\tilde G}\cdot\vectorize{\rho} \\
		&= \mathcal{E}\cdot\mathcal{G}\cdot\vectorize{\rho}
\end{align}
We shall work in the basis Pauli matrices, defining $\Sigma = \left\{I, \sigma_x, \sigma_y, \sigma_z\right\}^{\otimes n}$ as the collection of all $4^n$ $n$-qubit Pauli operators.  In this basis,
\begin{equation}
  \vectorize{\rho}_i = \tr\!\left(\rho \cdot\Sigma_i\right)/2^n,
\end{equation}
and
\begin{equation}
	(\mathcal{\tilde G})_{ij} = \tr\!\left(\mathsf{\tilde G}\left(\Sigma_j\right) \cdot\Sigma_i\right)/2^n.
\end{equation}
Process matrices written in Pauli basis are is often referred to as a \emph{Pauli transfer matrix} \cite{Chow2012}. Error maps take an especially nice form in this representation:
\begin{equation}\label{eq:process_matrix}
\mathcal{E} =
	\left(\begin{array}{c|cccc}
		1 &  & \vec{0}^T & \\ 
		\hline & &  &  \\
		\vec{m} &  & R &  \\
		 &  &  & 
	\end{array} 	
	\right)
\end{equation}
The top row of all trace-preserving (TP) maps is fixed to $\{1,0,0,0,\cdots\}$ and the remainder of the first column, $\vec{m}$, describes any deviations from unitality, as could arise from amplitude damping. If the error map is unitary, then the error is coherent, and the unital submatrix $R$ is perfectly antisymmetric, corresponding to a rotation of the generalized Bloch vector. Importantly, if $R$ is diagonal, then the error  is Pauli stochastic, with each diagonal entry corresponding to the probability that its associated Pauli error occurs in each application of the gate. 
% If R is symmetric but not diagonal, then the channel is still stochastic, but the random errors consist of correlated Pauli operators (such as $X+Y$). For a single qubit, this describes everything, but the situation can be slightly more complicated for more qubits.
